---
title: "Day 1 Part 2"
author: "Christopher Loan"
date: "February 20, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Section 0: Set seed

Set a seed at `022023`

```{r}

```

# Section 1: Fit Decision Tree to `ptitanic` Dataset

## Section 1.1: load packages

Load 6 packages: 

1. `tidyverse`
2. `rsample`
3. `rpart`
4. `rpart.plot`
5. `vip`
6. `pdp`

If you did not install these already, you'll have to install each of these with the `install.packages()` function. e.g., `install.packages('tidyverse')` etc.


```{r}

```

## Section 1.2: set up data

load the `ptitanic` data from the `rpart.plot` package. 

```{r}

```

Use `relevel()` to ensure `died` is the reference group in the `ptitanic` dataset.

```{r}

```

Use `rsample` to split the data into training and testing sets.

```{r}

```

## Section 1.3: Fit a Decision Tree

Predict `survived` with all other variables in the dataset

**Hint:** there are 2 ways to include all variables:

* One is simply listing all of them, separated with `+` symbols. 
* The other is a shortcut where you list the outcome, a tilde, then a period (`.`): `outcome_name ~ .`

```{r}

```


## Section 1.4: Explore Tree Structure 

### Section 1.4.1: Look at the structure

#### Section 1.4.1a

print the tree object (i.e., just run the name of the stored model)

```{r}

```

#### Section 1.4.1b

Use `rpart.plot()` to view the structure of decision the tree in a cleaner format.

```{r}

```

#### Section 1.4.2

Use `vip()` to visualize the variable importance plot for the decision tree.

```{r}

```

### Section 1.5: Evaluate Model Performance (unseen data)

Use your withheld testing data to determine the percent of new predictions that are correct.

```{r}

```

```{r}

```


When done, return your focus to the presentation
