---
title: "Workshop Data Guide"
author: "Christopher M. Loan"
date: "Februrary 20-22, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, message = FALSE)
library(tidyverse)
library(DT)
outcome_next_year <- 'se_xpd_totl_gd_zs'
```

```{r}
pstr <- 
  function(variable = NULL){
    # default returns written outcome
    if(is.null(variable)) {variable <- outcome_next_year}
       
    written_name <- 
      key %>% 
      filter(janitor_version == variable) %>% 
      pull(indicator_name)
    
    return(written_name)
  }
```

```{r}
vars_to_include <- 
  c('se_xpd_totl_gd_zs',
  'it_mlt_main',
  'nv_ind_totl_zs',
  'se_pre_enrr',
  'sh_dyn_1014',
  'bm_gsr_totl_cd',
  'sh_dyn_0509',
  'se_prm_enrl',
  'ne_con_totl_zs',
  'sl_tlf_totl_in'
)

key <- 
  read.csv(here::here('data/keys', 'key_impute.csv'))

dat <- 
  read.csv(
    here::here(
      'data', 
      'se_asia_imputed_world_bank.csv'
    )
  ) %>% 
  select(
    country_name, 
    year,
    outcome_next_year,
    all_of(vars_to_include)
  )
```

# {.tabset}

## Data Key

Here is a searchable key for the South East Asian World Bank Data.

<u>**DISCLAIMER** Keep in mind the `{missRanger}` package was used for imputation of missing data. See supplemental files for exact data handling</u> 

___


```{r}
key %>% 
  datatable(
    filter = 'top',
    caption = 'janitor_version is named after a package used to clean variable names; this indicates the naming convention used in the workshop'
  ) 
```


## Links to World Bank Data {.tabset}

Use the `General Links` tab to explore the World Bank's website and their homepage for indicators. 

The `All Indicators` tab has exact links for each of these indicators.

___


### General Links

Here are general links that may be useful: 

* [General](https://data.worldbank.org/)
* [Data / Indicators](https://data.worldbank.org/indicators)

### All Indicators

```{r}
key %>% 
  mutate(
    link = paste0('https://data.worldbank.org/indicators/', indicator_code)
  ) %>% 
  select(janitor_version, indicator_name, link) %>% 
  datatable(
    filter = 'top',
    caption = 'Links to Documentation'
  ) 
```


## Outcome Trajectory

As an outcome, we are predicting Next year's `r pstr()`.

___


```{r include = TRUE, fig.height=10, fig.width=15}
dat %>% 
  ggplot(
    aes(
      y = outcome_next_year, 
      x = year, 
      fill = country_name, 
      color = country_name
    )
  ) +
  labs(
    y = element_blank(), 
    title = paste0("Next Year's ", pstr(), ' over time, by country'),
    fill = 'Country\nName',
    color = 'Country\nName', 
    subtitle = paste0('Variable name = outcome_next_year')
  ) +
  geom_point(alpha = 0.5, size = 3) + 
  geom_smooth(se = FALSE) +
  theme_bw(base_size = 20) +
  theme(
    plot.title.position = 'plot',
    legend.position = 'bottom'
  ) +
  scale_x_continuous(
    breaks = seq(2000, 2020, 2)
  )
      
```


## DAY 3 SPOILERS: High Importance Features

Below are trajectories of items which have high variable importance from the tuned RF (end of day 3).

___


```{r include=TRUE, fig.height=10, fig.width=15}
vars_to_include %>% 
  walk(
    ~{
      p <- 
        dat %>% 
        ggplot(
          aes(
            y = !!sym(.x), 
            x = year, 
            fill = country_name, 
            color = country_name
          )
        ) +
        labs(
          y = element_blank(), 
          title = paste0(pstr(.x), ' over time, by country'),
          fill = 'Country\nName',
          color = 'Country\nName',
          subtitle = paste0('Variable name = ', .x)
        ) +
        geom_point(alpha = 0.5, size = 3) + 
        geom_smooth(se = FALSE) +
        theme_bw(base_size = 20) +
        theme(
          plot.title.position = 'plot',
          legend.position = 'bottom'
          ) +
        scale_x_continuous(
          breaks = seq(2000, 2020, 2)
        )
      
      cat('\n\n\n')
      print(p)
    }
  )

```

