---
title: "Day 1 Part 1 Lab Key"
author: "Christopher Loan"
date: "February 20, 2023"
output: html_document
---

\newpage 

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
```

# Set-up

## Installing packages

* Open `RStudio` by double clicking the `.Rproj` file

* Use `install.packages()` to install these packages:
    1. `tidyverse`
    2. `rsample`
    3. `rpart`
    4. `rpart.plot`
    5. `vip`
    6. `pdp`
    
```{r include=TRUE, eval=FALSE}
package_name_vector <- 
  c('tidyverse',
    'rsample',
    'rpart',
    'rpart.plot',
    'vip',
    'pdp'
  )

#install.packages(package_name_vector)
```

## Load Packages

* We will only use packages from `tidyverse` and `rpart.plot` for this example
* Use `library()` to load `tidyverse` and `rpart.plot`
* The other packages are in the examples at the end of the day

```{r}
library(tidyverse)
library(rpart.plot)
```


## Reading in Data

### From a package

Use `data()` to load the `ptitanic` data set (part of the `rpart.plot` package), then use `head(ptitanic, 4)` to see the first 4 rows


```{r}
data(ptitanic)
head(ptitanic, 4)
```

(we don't use this data anymore in this exercise, it was just for practice)

### From a .csv

Let's load 2 data files:

* Both are in a folder called `data`
* One is called `world_bank_key.csv` — store this object as `key`
* The other is called `se_asia_imputed_world_bank.csv`— store this object as `dat`

**Don't forget to store the data as an object with `<-`**

```{r}
key <- read.csv(here::here('data/keys', 'world_bank_key.csv'))
dat <- read.csv(here::here('data', 'se_asia_imputed_world_bank.csv'))
```

Here's a description of the files:

* `world_bank_key.csv` is a key which links the original variable name and indicator code, with a shorter name.
    * The shorter name is called `janitor_version` because I used a package called `janitor` to clean the `indicator_code` column.
    * the `janitor_version` correspond to the variable names (i.e., column names) in the `se_asia_imputed_world_bank.csv` file.
* `se_asia_imputed_world_bank.csv` 
    * yearly observations of several variables
    * variables with minimal missingness were imputed; see the cleaning data file provided in the OneDrive link

---

\newpage

# Cleaning Data

## Inspecting data

Convert both objects to tibbles and inspect them

```{r}
dat <- tibble(dat)
key <- tibble(key)
```

## Subsetting Data

run the following code to filter your key for only variables in this dataset:

**NOTE:** this key assumes you loaded the imputed data as `dat` and the key as `key`

```{r}
observed_names <- names(dat)

key_subset <- 
  key %>% 
  filter(
    janitor_version %in% observed_names
  )
```

**NOTE:** Because `%in%` starts with a `%`, you need back ticks (`), single quotes ('), or double quotes (") wrapping around it to call the help documentation. Like this: 
    
    ?`%in%`

## Inspecting Data

Use `dim()` with `key` and `key_subset` to see how much smaller this makes `key_subset`. Now it is managable to look through this.

```{r}
dim(key)
dim(key_subset)
```

\newpage

# Simple Refreshers

## Filtering

Use `key_subset` to find which variable in `dat` corresponds to an `indicator_name` of `Compulsory education, duration (years)`.

**HINT:** we want the one in the column labeled `janitor_version`

```{r}
key %>% 
  filter(
    indicator_name == 'Compulsory education, duration (years)'
  ) %>% 
  pull(janitor_version)
```

## Line Plots

Use `ggplot2` to plot the variable you found over time, and by country.

* in `ggplot(mapping = aes())`
    * set `year` as the `x`-axis 
    * set the variable corresponding to  `Compulsory education, duration (years)` on the `y`-axis.
    * set `country_name` as the `color`
* use `geom_line()` as our shape, since the values are continuous across time and are known.
* use `labs(y = 'Compulsory education, duration (years)')` in your plot to make a cleaner y-axis title

**HINT:** if your geom_line() is thinner than you'd like aesthetically, try `geom_line(size = 1.5)`. Don't go too large though, as it will make it harder to determine the value.

```{r}
plt <- 
  dat %>% 
  ggplot(
    aes(
      x = year, 
      y = se_com_durs, 
      color = country_name,
    )
  ) +
  geom_line(size = 1.5) +
  labs(y = 'Compulsory education, duration (years)')

plt
```

It is also very hard to see data for every country. 

## Continued Formating

formatting 

* let's divide this plot up by `country_name`
* this allows you to see each country's trajectory
* we do this by "facet wrapping", which makes the same plot within a group several times
* add this code to the end of your plot: ` + facet_wrap(vars(country_name))`

```{r}
plt + facet_wrap(vars(country_name))
```

\newpage

# Multi-Step Data Manipulations

(optional / if time allows)

## Exercise 1

### Subsetting Rows & Columns; Arranging

* Using `filter(year == 2018)`, subset the years of compulsory education in each country in 2018
* Use `select()` to limit the number of columns in the output to `country_name`, `year`, and `se_com_durs`
* Pipe (` %>% `) the filtered data to `arrange(desc(se_com_durs))` to show see which country has the highest current value 
* store this as an object called `compulsory_edu_2018`

```{r}
compulsory_edu_2018 <- 
  dat %>% 
  filter(year == 2018) %>% 
  select(country_name, year, se_com_durs) %>% 
  arrange(desc(se_com_durs)) 
```

* store `compulsory_edu_2018` as a `tibble()`.
* inspect that object 

```{r}
compulsory_edu_2018 <- tibble(compulsory_edu_2018)
compulsory_edu_2018
```

Which country had the highest `se_com_durs` in 2018?

## Exercise 2

### Filtering & Pulling

Find the variable which corresponds to `Government expenditure on education, total (% of GDP)`.

Last time: You could do this any way. Perhaps you inspected `key_subset` and found the variable. 
This time:

  * Get the `janitor_version` which corresponds to `Government expenditure on education, total (% of GDP)` with code
  * Do this by filling the arguments in this code: `key_subset %>% filter() %>% pull()` 

```{r}
key_subset %>% 
  filter(
    indicator_name == 'Government expenditure on education, total (% of GDP)'
  ) %>% 
  pull(janitor_version)
```

### Create a line plot of Government expenditure on education, total (% of GDP) vs. year

* Add points to your `geom_line()` as well
* Be sure to label the y axis with `labs(y = '% of GDP on education')`
* Use `?labs()` to figure out how to add a title. 
* Write your own title for the plot.

```{r}
dat %>% 
  ggplot(
    aes(
      x = year, 
      y = se_xpd_totl_gd_zs,
      fill = country_name,
      color = country_name
    )
  ) + 
  geom_point() +
  geom_line() + 
  labs(
    y = '% of GDP on education',
    title = 'Government expenditure on education (% GDP) over time'
    )
```

\newpage

# Diving Deeper

## Using Documentation to Understand Functions

Explain what this code is doing:

    dat %>% 
      group_by(country_name) %>% 
      summarize(
        mean_edu_expendature = mean(se_xpd_totl_gd_zs, na.rm = T)
      )

* Call help documentation if you don't know a variable name (e.g., `?summarize()`) 
* Try to figure it out from the help documentation before you run it
* Rewrite it in your own words here:


copy the code you just described, store the object as `education_expen`

```{r}
education_expen <- 
  dat %>% 
  group_by(country_name) %>% 
  summarize(
    mean_edu_expendature = mean(se_xpd_totl_gd_zs, na.rm = T)
  )
```

## More Visualization Practice

* Use a column plot `geom_col()` to visualize 
    * the `mean_edu_expendature` on the `x`-axis 
    * the `country_name` on the `y`-axis

```{r}
education_expen %>% 
  ggplot(
    aes(
      y = country_name,
      x = mean_edu_expendature
    )
  ) + 
  geom_col()
```

This doesn't look very nice, yet, but it can (see the "extra time" section)

I provide some formatting (and error bars) below:

\newpage

# Explorations (if extra time)

With any remaining time:

* Run the following code to make a nicer looking plot
* Explore the help documentation for functions/arguments you do not know  - e.g., `?fct_reorder()`
* Make your own tweaks to this plot, or make another

```{r}
dat %>% 
  group_by(country_name) %>% 
  summarize(
    mean_edu_expendature = mean(se_xpd_totl_gd_zs, na.rm = T), 
    sd = sd(se_xpd_totl_gd_zs, na.rm = T),
    se = sd/sqrt(n()), 
    lower_ci = mean_edu_expendature - 1.96*se, 
    upper_ci = mean_edu_expendature + 1.96*se 
  ) %>% 
 mutate(
    country_name_ordered = 
# fct_reorder allows us to order country name by a variable
# if we use the ordered version, the plot looks nicer
      fct_reorder(.f = country_name, .x = mean_edu_expendature)
  ) %>% 
   ggplot(
    aes(
      y = country_name_ordered,
      x = mean_edu_expendature,
      xmin = lower_ci, 
      xmax = upper_ci,
      fill = country_name
    )
  ) + 
  geom_errorbar(width = 0.4, color = 'black') + 
  geom_col(
    alpha = 0.8,
    color = 'black',
    show.legend = FALSE
  ) +
  theme_bw() +
  labs(
    y = element_blank(),
    x = '(%) of total GDP spent on education',
    fill = 'Country',
    title = 'Average Percentage (%) of GDP spent on Education',
    subtitle = paste0(min(dat$year), ' to ', max(dat$year)),
    caption = 'Source World Bank\n<https://data.worldbank.org/topic/education>\nError bars represent 95% CIs from t-distribution\nCritical value of 1.96'
    ) +
  theme(
    plot.title.position = 'plot',
    plot.caption.position = 'plot'
  )
```
