---
title: "Day 1\nPre-Workshop R Refresher"
author: "Christopher M. Loan"
date: 'February 20, 2022'
output: 
  xaringan::moon_reader:
    css: ["default"]
    nature:
      slideNumberFormat: "%current%/%total%"
---
layout: true
<div style="position: absolute;left:60px;bottom:11px;color:gray;">`r gsub('hristopher M', '', rmarkdown::metadata$author)`</div>
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(emojifont)
```

```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)

chk._ <- emoji('white_check_mark')
pkg._ <- emoji('package')
cmptr._ <- emoji('computer')
sml._ <- emoji('upside_down_face')
c._ <- emoji('car')
```

.right[**Setting Expectations**]

# Day 1 Agenda 

* Introduction and Overview **(9:00 am - 11:59 am)**
    * `r chk._` Introductions & KZ Lecture 
    * `r chk._` coffee break
    * `R` refresher (**up next**)
    * `ggplot2` refresher
    * Lab activity
  
* Lunch **(12:00 pm - 12:59 pm)**

* Describing and Fitting Decision Tree Models **(1:00 pm - 4:00 pm)**

---

.right[**Setting Expectations**]

# Communications

I want this to be a comfortable environment for all attendees. I'm a big fan of informal `r sml._` 

Raise your hand when you have a question; I'll answer at the end of my current thought. 

Do not wait until the end of the lecture, as you'll likely be more confused later (and I promise others have the same question).

If you are trying to follow along in `R` and have a technical problem that does not concern the group, please 

  1. keep following along conceptually, and 
  
  2. ask me for help during the break!

---

.right[**Setting Expectations**]

# Goals of the Workshop

Throughout the workshop, my goal is to

* demonstrate basic workflows for simple recursive partitioning models

* provide an <u>applied workshop</u>, which trades some depth for applied skills / coding

* build your comfortability enough that you do not feel lost in the peer-reviewed literature.

____

Throughout the workshop, I expect you to:

* ask questions when you're unsure

* tell me when I use unfamiliar language / statistical terms

* use the help documentation in `R` 

    * type `?example_function_name()`, or
    * use the RStudio interface (`help` tab located in the lower right panel)

---

.right[**Setting Expectations**]

# Intended Workshop Outcomes

I made these workshop materials to be a reference long-term. I expect you to refer to them to apply these methods. 

___

When applying these skills, I expect you will:

* refer to these slides to get started
* consult the the peer-reviewed literature before implementing these in research or industry
* continuously refer to `R` help documentation for these methods as needed

___

I don't expect you to

* follow along in `R`, unless you're comfortable doing so
* understand all of the things I do in `R` before you try them 
* be an expert in these methods at the end of the workshop

---

class: middle, center

# `R` Basics

---

.right[**`R` Basics**]

# Packages `r pkg._` in `R`

Like any software: `R` can only interact with objects, functions, data, etc. that you load into it.

You can make these yourself, but it is more common to load other people's functions, etc. 

To conduct analyses, we load other people's work - called "packages" - mostly from CRAN (Comprehensive R Archive Network)

You will also need to install several packages in this workshop. You can do that with:

* `install.packages()`
* the `RStudio` interface (`Packages tab on lower-right panel`)

---

.right[**`R` Basics**]

# Reading in Data `R` `r cmptr._`

#### Packaged Data

Data loaded from a package can be loaded with `data()`:

```{r}
library(rpart.plot)
ptitantic <- data(ptitanic)
```

#### Local Data

* if you have your own data, `R` needs to know the file path
* use `getwd()` to check your current file path

```{r}
getwd()
```

* Using projects (`.Rproj` files) simplifies file paths by searching within the project folder 

---

.right[**`R` Basics**]

# Import from `.csv`

Data loaded from a `.csv` can be loaded with `read.csv()`. 

```{r echo=TRUE, eval=FALSE, include=TRUE}
ex_dat <- read.csv('data/example_filename.csv')
```

Here, the `data` in the file name tells it that `example_file.csv` is in the `data` folder, separated by a slash.


#### Trying to get ahead of some potential errors

Last I knew, `R` had different defaults for which slash to use for mac (`/`) versus pc (`\`)

If some file paths do not work in my code:

* try switching the slash direction - (`/`) vs (`\`)

* use `file.path('folder', 'path')`

* ask for help

---

.right[**`R` Basics**]

# Workshop Data

This social science data reported by the World Bank. I provide extensive documentation of this in the supplemental materials.

Data is filtered to the years 2000 - 2018 & countries in South East Asia

* folder = `data`

* file = `se_asia_imputed_world_bank.csv`

```{r echo=TRUE, eval=FALSE, include=TRUE}
dat <- read.csv('data/se_asia_imputed_world_bank.csv')
```

```{r include = FALSE, echo = FALSE}
# doing it this way for xaringan
dat <- read.csv(here::here('data', 'se_asia_imputed_world_bank.csv'))
```

---

.right[**`R` Basics**]


# Inspecting data

```{r}
tibble(dat)
```

---

.right[**Data Types & Structures**]

# Inspecting data

It is `r nrow(dat)` observations (rows) of data.

Each observation (row) has many variables (columns) which correspond to it. 

A row represents an observation of a given country in a given year. 

---

class: center, middle

# Data Types & Structures

I am oversimplifying this section a little bit. Refer the [R-documentation](https://cran.r-project.org/doc/manuals/r-release/R-intro.html) for more details 

---

.right[**Data Types & Structures**]

# Data Types

`R` processes data differently based on its type

For this workshop, you'll need to understand some data types in order to work in `R`. 

* logical (dichotomous = TRUE/FALSE)
* numeric / double (numeric values with decimals)
* integer (whole numbers)
* characters / strings (words)


---

.right[**Data Types & Structures**]

# Vectors

**Definition** 

multiple observations of one data type e.g., `c(1, 2, 3)`

**Working with vectors**

  * access data in vector by its position or name (if named)
  * example: 3 run times (in seconds) for 3 different complexity models
  
```{r}
run_time_seconds <- 
  c('simple model' = 30, 
    'intermediate model' = 58, 
    'complex model' = 90)
# by position
run_time_seconds[1]
# by name
run_time_seconds['simple model']
```

---

.right[**Data Types & Structures**]

.left[### Lists]

.pull-left[

```{r include = TRUE, echo = TRUE}
example_list <- 
  list(
    model = 
      lm(data = dat, 
         formula = se_xpd_totl_gd_zs ~ year + country_name),
    plot = 
      ggplot(
        data = dat, 
        aes(se_xpd_totl_gd_zs, year)
        ) + 
      geom_smooth(se = FALSE, show.legend = FALSE)
  )
```

```{r include = TRUE, echo = TRUE, fig.width=6, fig.height=2}
example_list[['plot']]
```

]

.pull-right[

(for our purposes) **a list is a vector that can hold complex data**
  
  * we will use this to store models, plots, etc.
  * Refer to the object in the list with double brackets


```{r include = TRUE, echo = TRUE, fig.width=6, fig.height=2}
example_list[[2]]
```

]


---

.right[**Data Types & Structures**]

.left[

# Data Frames


  * each cell is coded as [row_number, column_number]
  * each column can be its own data type, but must be constant within column
  * Refer to variables with `$` and name, or by position
  
  
]
  
.pull-left[
  
```{r,fig.width=1}
iris$Species
```

]
.pull-right[

```{r fig.width=1}
iris[,5]
```
]

---

.right[**Data Types & Structures**]


# Factors

  * a method of storing categorical data (i.e., convert a vector to a factor)
  
  * factors are typically unordered categories (e.g., Ocean names, brands), made by converting strings / characters to factors
  
  * factors can be ordered too, though (e.g., Likert- / Rating-Scale; ordered factors won't be relevant for our workshop).
  
---

.right[**Data Types & Structures**]


# Factors

Setting variables as factors is easy:

```{r}
class(dat$country_name)
dat$country_name <- factor(dat$country_name)
class(dat$country_name)
```

---

class: center, middle

# `r emoji('bar_chart')` ggplot2 refresher

(because this workshop is very visual)

---

.right[**`r emoji('bar_chart')` ggplot2 refresher**]

# Logic

* observed data is mapped to several "aesthetics" 
    * `x-axis`, `y-axis`,`group`, `color`, `fill`, `size`, etc.
    * this concept is easiest to understand with an `x-` and `y-axis`
    
* Layers which represent those aesthetics are layered on top
    * `geom_point` = scatter plot
    * `geom_smooth` = line of best fit between `y` and `x` aesthetics
    * `geom_col` = column plot
    
* `ggplot2` objects are basically stacked in layers — add layers with the `+` symbol

---

.right[**`r emoji('bar_chart')` ggplot2 refresher**]

# Mapping Aesthetics

* y-axis = `se_xpd_totl_gd_zs` (Government expenditure on education as percent of GDP)  
* x-axis = `year`

`ggplot2` will use data to choose limits to these, but no data will appear

```{r}
plot_mapping <- 
  ggplot(
    data = dat, 
    mapping = 
      aes(
        y = se_xpd_totl_gd_zs, 
        x = year
      )
  )
```


---

```{r fig.height=8, fig.width=14}
plot_mapping
```


---

# `geom_point()`


```{r fig.height=8, fig.width=14}
plot_mapping + geom_point()
```

---

# `geom_smooth()`

```{r fig.height=8, fig.width=14}
plot_mapping + geom_point() + geom_smooth()
```


---

### `geom_smooth()` by `country_name`

```{r fig.height=8, fig.width=14}
plot_mapping + geom_point() + geom_smooth() + 
  aes(color = country_name, fill = country_name) 
```

---

# Format the plot

```{r}
cleaner_plot <- 
  plot_mapping + 
  aes(color = country_name, fill = country_name) +
  geom_point() + 
  geom_smooth() +
  theme_bw(base_size = 15) + 
  scale_y_continuous(
    breaks = seq(0, 15, 3), 
    limits = c(0, 15)
  ) +
  scale_x_continuous(
    breaks = seq(2000, 2018, 3)
  ) +
  theme(
    plot.title.position = 'plot',
    plot.caption.position = 'plot',
    legend.position = c(0.5, 0.8),
    legend.direction = 'horizontal'
    )
```

---

# Add labels

```{r}
plt_labled <- 
  cleaner_plot + 
  labs(
    y = 'Government Expendature on Education (% of GDP)',
    caption = 'World Bank Education Data: 2000:2018',
    title = 'Trajectories of Government Expendature on Education (% of GDP) by Country',
    color = 'Country',
    fill = 'Country'
    ) 
```

---

```{r echo=FALSE, include=TRUE, fig.align='center', fig.width=10, fig.height=8}
plt_labled
```

---
class: center, middle

# [Exit Slides & Give `?`/Documentation Tutorial]

---
class: center, middle

# Activity
